---
output: github_document
---


- Define a model
- Define action space. Under the assumption of fixed action, problem is not MDP and trivial to solve by sweep through possible fixed actions.

```{r}
library(tidyverse)
```



## Five species

```{r}

model_5sp <- function(u_prey = 0.5,
                      u_pred = 0.5,
                      sigma = 0.05,
                      x1o = 0.44, # unfushed equib
                      x2o = 0.17,
                      x3o = 0.186,
                      x4o = 0.600,
                      x5o = 0.280,
                      Tmax = 50) {
    x1 <- x2 <- x3 <- x4 <- x5 <- numeric(Tmax)
    s1 <- s2 <- s3 <- s4 <- numeric(Tmax)
    xi1 <- rnorm(Tmax, 0, sigma)
    xi2 <- rnorm(Tmax, 0, sigma)
    xi3 <- rnorm(Tmax, 0, sigma)
    xi4 <- rnorm(Tmax, 0, sigma)
    xi5 <- rnorm(Tmax, 0, sigma)

    x1[1] <- x1o
    x2[1] <- x2o
    x3[1] <- x3o
    x4[1] <- x4o
    x5[1] <- x5o
   
    for (t in 1:(Tmax-1)) {
        s1[t] <- x1[t] * (1 - u_prey)
        s2[t] <- x2[t] * (1 - u_prey)
        s3[t] <- x3[t] * (1 - u_prey)
        s4[t] <- x4[t] * (1 - u_pred)

        x1[t+1] = s1[t]*exp(1.0213 - s1[t] - 0.0861*s2[t] - 0.3141*s3[t] - 0.7252*s4[t] - 0.2445*x5[t] + xi1[t])
        x2[t+1] = s2[t]*exp(1.0289 - 0.4765*s1[t] - s2[t] - 0.1370*s3[t] - 0.9811*s4[t] - 0.0915*x5[t] + xi2[t])
        x3[t+1] = s3[t]*exp(1.0207 - 0.3193*s1[t] - 0.3461*s2[t] - s3[t] - 0.6367*s4[t] - 0.8716*x5[t] + xi3[t])
        x4[t+1] = s4[t]*exp(0.7252*s1[t] + 0.9811*s2[t] + 0.6367*s3[t] - s4[t] + xi4[t])
        x5[t+1] = x5[t]*exp(0.2445*s1[t] + 0.0915*s2[t] + 0.8716*s3[t] - x5[t] +  xi5[t])

    }
    df <- tibble(t = 1:Tmax, x1, x2, x3, x4, x5)
    df
}

```

```{r}
clip <- function(x, lower=0, upper=1) {
    x[x <= lower] <- lower
    x[x >= upper] <- upper
    x
}

utility <- function(x) {
    delta <- 0.999
    w3 <- 0.5
    w2 <- 0.25
    w1 <- 1 - (w2 + w3)
    u_prey <- clip(x[1], 0, 1)
    u_pred <- clip(x[2], 0, 1)
    df <- model_5sp(u_prey, u_pred)
    R <- 
        w1 * u_prey * (df$x1 + df$x2 + df$x3) +
        w2 * u_pred * df$x4 +
        w3 * df$x5
    t <- seq_along(R)
    - as.numeric(R %*% delta ^ t)
    }

-utility(c(0.5,0.5))
-utility(c(0,1))
-utility(c(1,0))
-utility(c(0,0))

o <- optim(c(0.,0.), utility, method="L-BFGS-B", lower = c(0,0), upper = c(1,1))
o$par
o$value

o <- optim(c(0.5,0.1), utility)
clip(o$par, 0, 1)
o$value

o <- optim(c(0.5,0.5), utility, method="SANN")
clip(o$par, 0, 1)
o$value

```

For the sake of argument, let us allow the model of Brias & Munch to represent the "true" model, which is unknown.
As researchers, we only have access to data derived from observations from the model. 
Let us consider a sample simulation of such data:

```{r}
df <- model_5sp(0, .6, sigma = 0.01)

df %>%
    pivot_longer(tidyselect::starts_with("x"),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(t, biomass, col= species)) + geom_path()
```

There are many other models, including process-based models, statistical models or even black-box machine learning models
which 5 dimensional state-space and two-dimensional action space. 



```{r}
df <- model_5sp(0, 0., sigma = 0.01)

df %>%
    transmute(time = t, cormorant = x5, bass = x4, herring = x1+x2+x3) %>%
    pivot_longer(any_of(c("cormorant", "bass", "herring")),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(time, biomass, col= species)) + geom_path()
```

A reasonable single-species model for Bass achieves the optimal solution.
The model is reasonable in the sense that it is a simple Ricker model 
that comparably reproduces the unfished dynamics of the Bass.
Nevertheless, it is actually quite poor at forecasting fished stock sizes.

```{r}

model_1sp <- function(u = 0, A = 2, R = 1.6, Tmax=50, x1o = .8, sigma = 0.01) {

    omega1 <- rnorm(Tmax, 0, sigma)

    s1 <- x1 <- numeric(Tmax)
    x1[1] <- x1o

    for (t in 1:(Tmax-1)) {
        s1[t] <- x1[t] * (1 - u)
        x1[t+1] <- s1[t] * exp(R - s1[t] * A + omega1[t])
    }

    df <- tibble(time = 1:Tmax, x1)
}

df <- model_1sp(0)
df %>% ggplot(aes(t, x1)) + geom_path()

```


```{r}
utility <- function(x) {
    delta <- 0.999
    u <- clip(x, 0, 1)
    df <- model_1sp(u)
    R <- u * df$x1 
    t <- seq_along(R)
    - as.numeric(R %*% delta ^ t)
    }

o <- optimize(utility, c(0,1))
o$minimum
```

```{r}
df <- model_1sp(o$minimum)
df %>% ggplot(aes(t, x1)) + geom_path()
```



```{r}
R1 <- R2 <- 3
A1 <- A2 <- 2
B1 <- 0.5
B2 <- -0.3
Tmax <- 50
omega1 <- rnorm(Tmax, 0, 0.001)
omega2 <- rnorm(Tmax, 0, 0.001)

u1 <- u2 <- s1 <- s2 <- x1 <- x2 <- numeric(Tmax)
x1[1] <- 1
x2[1] <- 1

# or, constant strategy
u1 <- rep(.2, Tmax)
u2 <- rep(.5, Tmax)

for (t in 1:(Tmax-1)) {
    s1[t] <- x1[t] * (1 - u1[t])
    s2[t] <- x2[t] * (1 - u2[t])
    x1[t+1] <- s1[t] * exp(R1 - A1 * s1[t] - B1 * s2[t] + omega1[t])
    x2[t+1] <- s2[t] * exp(R2 - A2 * s1[t] - B2 * s1[t] + omega2[t])
}

df <- data.frame(time = 1:Tmax, x1, x2, u1, u2)
```



